Debugging by \textit{lastChange}

### 2010-10-21 papers/pldi2011/sigplanconf-template.tex

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
We implemented \textit{Querypointer}, an Extension to Firebug JavaScript debugger, which provides \textit{lastChange} functionality. Firebug is an add-on of Firefox browser. Firefox JavaScript engine provides a JavaScript debugging interface and \textit{Querypointer} is developed over this interface. The debugging support in JavaScript comparing to more traditional languages such as Java and C\# is less comprehensive. Some part of this flaw is due to the dynamic nature of JavaScript. For example, JavaScript does not have classes and therefore setting watchpoint on properties before program execution is not possible. We explain how we could overcome these limitations and provide required operations for \textit{lastChange} algorithm implementation. Moreover, we breifly describe mechanisms for automated execution reproduction, data collection and presentation.

\subsection{Algorithm technical details}
\subsubsection{\texttt{objectId()}}
The object id is kept as a property of the object, \texttt{\_objectid}, which is not enumerable. Therefore, it will not appear in \texttt{for(property in object)} loops and has no effect on the executing program. Once \texttt{objectId()} is called on an object and the object has not \texttt{\_objectId} property, this property is defined by calling \texttt{defineProperty} JavaScript function. As a parameter it can be specified that the defined property is not enumerable. Then it is enough to assign a new generated id to this property.

\subsubsection{\texttt{scopeId()}}
The scope id is kept as a variable in the scope, \texttt{\_scopeId}. As we said, a hook is set on the first instruction of the function defines the variable. Whenever this hook hits,  \texttt{\_scopeId} variable is set by \texttt{eval()} function. For example evaluating \texttt{var \_scopeId = 50} will create a variable with name \texttt{\_scopeId} and value \texttt{50} in the local scope.

\subsubsection{\texttt{setPropertyChangeHook()}}
To implement this operation, we use two features that are available in Firefox JavaScript engine. First, every regular JavaScript object has a \texttt{watch()} function which receives two parameters, a property name and another function as a callback. Whenever the property with the given name changes the calleback function is called. The hook is set by this function is enabled even if the property is removed and added again [ref]. The issue with this function is that it can only be applied to available objects. However, at the beginning of re-execution there is no object available. To solve this issue we use the second feature. Enabling a flag gives us the place the object created including the JavaScript file url and line number (e.g., myFile.js, line 24). Though this is a limited data, it helps us to get a reference to the object once it is created. Then it is enough to set objet id and call \texttt{watch()} with the property name. Consider that in this way, we implicitly assume that the object will be created at the same place in the next re-executions. This assumption will be discussed later.

In the next section, we explain the algorithm for getting a reference to the object after its creation. We could replace this algorithm with a much simpler algorithm if Firefox supported these two features. First instead of line number of object creation point, it gave us the program counter. Second, there was a way to hook object creation and get a reference to the created object. Having this two features, it was enough to set a hook at the instruction and once the execution reaches to instruction, we enable the object creation hook and we step one instruction and we get the reference to the object.

>>>>

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
We implemented \textit{Querypointer}, an Extension to Firebug JavaScript debugger, which provides \textit{lastChange} functionality. Firebug is an add-on of Firefox browser. Firefox JavaScript engine provides a JavaScript debugging interface and \textit{Querypointer} is developed over this interface. The debugging support in JavaScript comparing to more traditional languages such as Java and C\# is less evolved. Some part of this flaw is due to the dynamic nature of JavaScript. For example, JavaScript does not have classes and therefore setting watchpoint on properties before program execution is not possible. We explain how we could overcome these limitations and provide required operations for \textit{lastChange} algorithm implementation. Moreover, we breifly describe mechanisms for automated execution reproduction, data collection and presentation.

\subsection{Algorithm technical details}
\subsubsection{\texttt{objectId()} operation}
\texttt{objectId()} operation receives a regular JavaScript object and returns an id which is unique during one execution. This operation keeps the object id in an object property with name \texttt{\_objectId}. Once this operation receives an object, it returns \texttt{\_objectId} value if this property is already defined, otherwise it generates a new id and sets this property object. To set the property, it calls \texttt{defineProperty} standard JavaScript function. \texttt{defineProperty} receives an argument which specifies whether the property is enumerable or not. By setting the enumerable field \texttt{false} for \texttt{\_objectId}, this property will not apear in \texttt{for(property in object)} loops and therfore has no effect on the program execution. Consider that \texttt{\_objectId} is not set for all objects but only those objects that need an id.

\subsubsection{\texttt{scopeId()} operation}
The scope id is kept as a variable with name \texttt{\_scopeId} in the scope. We explained that we set a hook at the first instruction of the function which defines the variable. Once this hook hits, \texttt{\_scopeId} is set by calling \texttt{eval()} function. For example, executing \texttt{eval( var \_scopeId = 10)} creates a variable with name \texttt{\_scopeId} and value \texttt{10} in the local scope which is our interesting scope.

\subsubsection{\texttt{setPropertyChangeHook()} operation}
We implement a slightly different version of this operation. Setting a flag into the Firefox JavaScript engine we can get the place am object created including the JavaScript file url and line number (e.g., myFile.js, line 24). We can assume that an object is created in the same place in the next re-executions. This assumption does not lead to a wrong answer. We will discuss it in the next section when we discuss the effect of non-determinism. Having this assumption, it is enough to set property change hook for the objects created at one place.

The Firefox JavaScript engine supports watching property changes in an object. Every object has a \texttt{watch()} function which receives two parameters, a property name and another function as the callback. Whenever the property with the given name changes the calleback function is called. The hook is set by this function remains enabled even if the property is deleted and defined again. This function can only be applied to avaialbe objects. However, at the beginning of execution, no object excepting the global object is available.

If we could get a reference to the object just after its creation, we could use \texttt{watch()} function to monitor property changes in the object. In the next section we explain the algorithm we devised for this purpose by parsing and analyzing the code. This algorithm could be replaced by a much simpler one if Firefox debugging interface gives us the program counter of object creation instead of line number and also provides object creation hook.
By these assumptions, it was enough to set a hook on the object creation location. Once the execution reaches this location, by enabling object creation hook, we could get a reference to the new object.
